---
title: "Applying mcSimulation function in decisionSupport"
author: "Cory Whitney, Eduardo Fernandez, and Eike Luedeling <br /><small>INRES-Horticultural Sciences, University of Bonn, Auf dem Huegel 6, 53121 Bonn, Germany<small/>"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Applying mcSimulation function in decisionSupport}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
#bibliography:
 # - packages.bib
  #- wildfire.bib 
#csl: elsevier-harvard.csl
---
<!-- <style> -->
<!-- img { -->
<!-- display: block; -->
<!-- margin: 1!important; -->
<!-- padding: 0!important; -->
<!-- border: 0!important} -->
<!-- </style> -->

<!-- <img src = "sticker.png" alt = "decisionSupport logo" align = "right" width = "20%" height = "20%"/> -->

<style>
body {
text-align: justify}
</style>

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, warning = F, include = F}
#Automatically write R package citation entries to a .bib file
knitr::write_bib(c(.packages(), 
                   'chillR',
                   'dplyr',
                   'ggplot2', 
                   'kableExtra',
                   'decisionSupport'), 'packages.bib')
require(tidyverse)
require(kableExtra)
library(decisionSupport)
```

## Introduction

In this vignette we use the libraries from the tidyverse collection. Install the libraries from CRAN using the `install.packages()` command. For using the libraries, load the packages with the `library()` command.

```{r install_packages, eval = FALSE}
install.packages("tidyverse")

library(tidyverse)
```

Here we generate a model as a function using `decisionSupport` library. We applied the functions `vv()` which produces time series with variation from a pre-defined mean and coefficient of variation, `chance_event()` which randomly simulate whether events occur, and `discount()` which discounts values along a time series.

```{r model}
example_decision_function <- function(x, varnames){
  # ex-post risk: impacts the benefits ####
  # None included yet
  # ex-ante risk: impacts the implementation of interventions ####
  intervention_NonPopInvolvEvent <-
    chance_event(intervention_NonPopInvolv, 1, 0, n = 1)
  
  # calculation of common random draws for all intervention model runs ####
  
  TLU <- vv(TLU_no_intervention, var_CV, n_years)
  TLU_profit <- vv(profit_per_TLU, var_CV, n_years)
  
  precalc_intervention_fruit_benefits <-
    vv(intervention_fruit_area_ha, var_CV, n_years) *
    vv(intervention_fruit_yield_t_ha, var_CV, n_years) *
    vv(intervention_fruit_profit_USD_t, var_CV, n_years)
  
  precalc_intervention_vegetable_benefits <-
    vv(intervention_vegetable_area_ha, var_CV, n_years) *
    vv(intervention_vegetable_yield_t_ha, var_CV, n_years) *
    vv(intervention_vegetable_profit_USD_t, var_CV, n_years)
  
  precalc_intervention_rainfed_crop_benefits <-
    vv(intervention_rainfed_crop_area_ha, var_CV, n_years) *
    vv(intervention_rainfed_crop_yield_t_ha, var_CV, n_years) *
    vv(intervention_rainfed_crop_profit_USD_t, var_CV, n_years)
  
  #  Intervention ####
  
  for (decision_intervention_strips in c(FALSE,TRUE))
      {
      
  if (decision_intervention_strips)
  {
    intervention_strips <- TRUE
    intervention_strips_PlanningCost <- TRUE
    intervention_strips_cost <- TRUE
  } else
  {
    intervention_strips <- FALSE
    intervention_strips_PlanningCost <- FALSE
    intervention_strips_cost <- FALSE
  }
  
  if (intervention_NonPopInvolvEvent) {
    intervention_strips <- FALSE
    intervention_strips_cost <- FALSE
  }
  
  # Costs ####
  if (intervention_strips_cost) {
    cost_intervention_strips <-
      intervention_adaptation_cost + intervention_tech_devices_cost + intervention_nursery_cost +
      intervention_wells_cost +
      intervention_training_cost + intervention_mngmt_oprt_cost + intervention_mngmt_follow_cost +
      intervention_mngmt_audit_cost
  } else
    cost_intervention_strips <- 0
  
  if (intervention_strips_PlanningCost) {
    plan_cost_intervention_strips <-
      intervention_communication_cost + intervention_zoning_cost
  } else
    plan_cost_intervention_strips <- 0
  
  maintenance_cost <- rep(0, n_years)
  
  if (intervention_strips)
    maintenance_cost <-
    maintenance_cost + vv(maintenance_intervention_strips, var_CV, n_years)
  
  intervention_cost <- maintenance_cost
  intervention_cost[1] <-
    intervention_cost[1] + cost_intervention_strips + plan_cost_intervention_strips

  
  # Benefits from  cultivation in the intervention strips ####
  
  intervention_fruit_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_fruit_benefits
  intervention_vegetable_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_vegetable_benefits
  intervention_rainfed_crop_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_rainfed_crop_benefits
  
  # Total benefits from crop production (agricultural development and riparian zone) ####
  crop_production <-
    intervention_fruit_benefits +
    intervention_vegetable_benefits +
    intervention_rainfed_crop_benefits
  
  # Benefits from livestock ####
  # The following allows considering that intervention strips may
  # restrict access to the reservoir for livestock.
  
  if (intervention_strips)
    TLU_intervention <-
    TLU * (1 + change_TLU_intervention_perc / 100)
  else
    TLU_intervention <- TLU
  
  if (decision_intervention_strips){
  
    livestock_benefits <- TLU_intervention * TLU_profit
    
    total_benefits <- crop_production + livestock_benefits
    
    net_benefits <- total_benefits - intervention_cost
    
    result_interv <- net_benefits}
  
  
  if (!decision_intervention_strips){
    
    livestock_benefits <- TLU_no_intervention * TLU_profit
    
    total_benefits <- livestock_benefits
    
    net_benefits <- total_benefits - intervention_cost
    
    result_n_interv <- net_benefits}
  
    } #close intervention loop bracket

NPV_interv <-
  discount(result_interv, discount_rate, calculate_NPV = TRUE)

NPV_n_interv <-
  discount(result_n_interv, discount_rate, calculate_NPV = TRUE)

# Beware, if you do not name your outputs (left-hand side of the equal sign) in the return section, the variables will be called output_1, _2, etc.

return(list(NPV_decision_do = NPV_interv - NPV_n_interv,
            Cashflow_decision_do = result_interv - result_n_interv))
}

```

Using the model above, we can perform a Monte Carlo simulation with the `mcSimulation()` function from `decisionSupport`. This function generates a distribution representing the desired outputs (see `return()` function above) by calculating random draws in our defined `example_decision_function()`. 

```{r mcSimulation}
test_mcSimulation_function <- decisionSupport::mcSimulation(
  estimate = decisionSupport::estimate_read_csv("example_input_table.csv"),
  model_function = example_decision_function,
  numberOfModelRuns = 1e4, #run 10,000 times
  functionSyntax = "plainNames"
)

```


```{r}
decisionSupport::plot_distributions(mcSimulation_object = test_mcSimulation_function, vars = c("NPV_decision_do"),
                   method = 'boxplot_density')
```


## References
