---
title: "Running a Monte Carlo in decisionSupport with the mcSimulation function"
author: "Cory Whitney, Eduardo Fernandez, and Eike Luedeling <br /><small>INRES-Horticultural Sciences, University of Bonn, Auf dem Huegel 6, 53121 Bonn, Germany<small/>"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Applying mcSimulation function in decisionSupport}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography:
 - packages.bib
 - vignettes.bib 
---


<style>
body {
text-align: justify}
</style>

```{r, include = F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, warning = F, include = F}
#Automatically write R package citation entries to a .bib file
knitr::write_bib(c(.packages(), 
                   'chillR',
                   'dplyr',
                   'ggplot2', 
                   'kableExtra',
                   'decisionSupport'), 'packages.bib')
require(tidyverse)
require(kableExtra)
library(decisionSupport)
```

### Decision analysis with the decisionSupport package

#### The decisionSupport package

Many of the variables decision makers need to consider in development cannot be precisely quantified, at least not without unreasonable effort. The major objective of (prescriptive) decision analysis is to support the decision-making processes when faced with this problem [@Luedeling2016]. Decision analysis can make forecasts of decision outcomes without precise numbers. 

The decisionSupport package [@R-decisionSupport] implements this as a Monte Carlo simulation, which generates a large number of plausible system outcomes, based on random numbers for each input variable, drawn from user-specified probability distributions. 

1)	an R function that predicts decision outcomes based on the variables named in a separate data table. This R function is customized by the user to address a particular decision problem to provide the decision analysis model.  
2)	an input table (in .csv format) specifying the names and probability distributions for all variables used in the decision model. These distributions aim to represent the full range of possible values for each component of the model.  

These two inputs are provided as arguments to the `mcSimulation` function, which conducts a Monte Carlo analysis with repeated model runs based on probability distributions for all uncertain variables. The data table and model are customized to fit the particulars of a specific decision.

#### Example – Reservoir Sedimentation in the Upper Volta River Basin of Burkina Faso

We showcase a partial example of a Monte-Carlo-based selection of a sedimentation management strategy for a reservoir in the Upper Volta River Basin of Burkina Faso [@Lanzanova2019]. These reservoirs have multiple benefits for rural communities and are important for food security and livelihoods. Sedimentation is a major impediment for the functioning of these reservoirs. The design of an efficient sedimentation management intervention requires assessment of multiple uncertain quantities and risks. Decision analysis is highly applicable in such data-scarce environments, where science has traditionally struggled to provide actionable information to policy-makers, development practitioners, NGOs and rural communities. 

Here we walk through a simple example of the use of various decisionSupport functions with the `tidyverse` libraries including `ggplot2`, `plyr` and `dplyr` among others in the [R programming language](https://www.r-project.org/) [@R-base]. 

Install the libraries from CRAN using the `install.packages()` command. For using the libraries, load the packages with the `library()` command.

```{r install_packages, eval = FALSE}
install.packages("tidyverse")

library(tidyverse)
```

Here we generate a model as a function using `decisionSupport` library. The example is a simple version of the model applied in Burkina Faso [@Lanzanova].  We use the `decisionSupport` functions `vv()` to produce time series with variation from a pre-defined mean and coefficient of variation, `chance_event()` to simulate whether events occur and `discount()` to discount values along a time series.

```{r model}
example_decision_function <- function(x, varnames){
  
  # calculate ex-ante risks: impact the implementation of interventions ####
  intervention_NonPopInvolvEvent <-
    chance_event(intervention_NonPopInvolv, 1, 0, n = 1)
  
  # pre-calculation of common random draws for all intervention model runs ####
  
  # profits from Tropical Livestock Units (TLU)
  TLU <- vv(TLU_no_intervention, var_CV, n_years)
  TLU_profit <- vv(profit_per_TLU, var_CV, n_years)
  
  # benefits of fruit
  precalc_intervention_fruit_benefits <-
    vv(intervention_fruit_area_ha, var_CV, n_years) *
    vv(intervention_fruit_yield_t_ha, var_CV, n_years) *
    vv(intervention_fruit_profit_USD_t, var_CV, n_years)
  
  # benefits of vegetables
  precalc_intervention_vegetable_benefits <-
    vv(intervention_vegetable_area_ha, var_CV, n_years) *
    vv(intervention_vegetable_yield_t_ha, var_CV, n_years) *
    vv(intervention_vegetable_profit_USD_t, var_CV, n_years)
  
  # benefits of rainfed crops
  precalc_intervention_rainfed_crop_benefits <-
    vv(intervention_rainfed_crop_area_ha, var_CV, n_years) *
    vv(intervention_rainfed_crop_yield_t_ha, var_CV, n_years) *
    vv(intervention_rainfed_crop_profit_USD_t, var_CV, n_years)
  
  #  Intervention ####
  
  for (decision_intervention_strips in c(FALSE,TRUE))
      {
      
  if (decision_intervention_strips)
  {
    intervention_strips <- TRUE
    intervention_strips_PlanningCost <- TRUE
    intervention_strips_cost <- TRUE
  } else
  {
    intervention_strips <- FALSE
    intervention_strips_PlanningCost <- FALSE
    intervention_strips_cost <- FALSE
  }
  
  if (intervention_NonPopInvolvEvent) {
    intervention_strips <- FALSE
    intervention_strips_cost <- FALSE
  }
  
  # Costs ####
  if (intervention_strips_cost) {
    cost_intervention_strips <-
      intervention_adaptation_cost + intervention_tech_devices_cost + intervention_nursery_cost +
      intervention_wells_cost +
      intervention_training_cost + intervention_mngmt_oprt_cost + intervention_mngmt_follow_cost +
      intervention_mngmt_audit_cost
  } else
    cost_intervention_strips <- 0
  
  if (intervention_strips_PlanningCost) {
    plan_cost_intervention_strips <-
      intervention_communication_cost + intervention_zoning_cost
  } else
    plan_cost_intervention_strips <- 0
  
  maintenance_cost <- rep(0, n_years)
  
  if (intervention_strips)
    maintenance_cost <-
    maintenance_cost + vv(maintenance_intervention_strips, var_CV, n_years)
  
  intervention_cost <- maintenance_cost
  intervention_cost[1] <-
    intervention_cost[1] + cost_intervention_strips + plan_cost_intervention_strips

  
  # Benefits from  cultivation in the intervention strips ####
  
  intervention_fruit_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_fruit_benefits
  intervention_vegetable_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_vegetable_benefits
  intervention_rainfed_crop_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_rainfed_crop_benefits
  
  # Total benefits from crop production (agricultural development and riparian zone) ####
  crop_production <-
    intervention_fruit_benefits +
    intervention_vegetable_benefits +
    intervention_rainfed_crop_benefits
  
  # Benefits from livestock ####
  # The following allows considering that intervention strips may
  # restrict access to the reservoir for livestock.
  
  if (intervention_strips)
    TLU_intervention <-
    TLU * (1 + change_TLU_intervention_perc / 100)
  else
    TLU_intervention <- TLU
  
  if (decision_intervention_strips){
  
    livestock_benefits <- TLU_intervention * TLU_profit
    
    total_benefits <- crop_production + livestock_benefits
    
    net_benefits <- total_benefits - intervention_cost
    
    result_interv <- net_benefits}
  
  
  if (!decision_intervention_strips){
    
    livestock_benefits <- TLU_no_intervention * TLU_profit
    
    total_benefits <- livestock_benefits
    
    net_benefits <- total_benefits - intervention_cost
    
    result_n_interv <- net_benefits}
  
    } #close intervention loop bracket

NPV_interv <-
  discount(result_interv, discount_rate, calculate_NPV = TRUE)

NPV_n_interv <-
  discount(result_n_interv, discount_rate, calculate_NPV = TRUE)

# Beware, if you do not name your outputs (left-hand side of the equal sign) in the return section, the variables will be called output_1, _2, etc.

return(list(Interv_NPV = NPV_interv,
            NO_Interv_NPV = NPV_n_interv,
            NPV_decision_do = NPV_interv - NPV_n_interv,
            Cashflow_decision_do = result_interv - result_n_interv))
}

```

Using the model function above, we can perform a Monte Carlo simulation with the `mcSimulation()` function from `decisionSupport`. This function generates a distribution representing the desired outputs (see `return()` function above) by calculating random draws in our defined `example_decision_function()`. 

```{r mcSimulation}
test_mcSimulation_function <- decisionSupport::mcSimulation(
  estimate = decisionSupport::estimate_read_csv("example_input_table.csv"),
  model_function = example_decision_function,
  numberOfModelRuns = 1e4, #run 10,000 times
  functionSyntax = "plainNames"
)

```

We can use the `plot_distributions()` function to produce one of the several plotting options for distribution outputs. This shows us an overlay of the full results of the Monte Carlo model of the decision options, i.e. the expected NPV if we choose to do the intervention `Interv_NPV` or not do the intervention `NO_Interv_NPV`.

```{r}
decisionSupport::plot_distributions(mcSimulation_object = test_mcSimulation_function, 
                                    vars = c("Interv_NPV", "NO_Interv_NPV"),
                                    method = 'smooth_simple_overlay')
```

We can use the same function for the value of the decision (difference in NPV between do and do not do). This is more helpful for us since it shows us the outcome distribution of the decision itself. 

```{r}
decisionSupport::plot_distributions(mcSimulation_object = test_mcSimulation_function, 
                                    vars = "NPV_decision_do",
                                    method = 'boxplot_density')
```

```{r}
plot_cashflow(mcSimulation_object = test_mcSimulation_function, cashflow_var_name = "Cashflow_decision_do")


```



## Work in progress (July 02, 2020)

Here we apply a post-hoc analysis to the mcSimulation outputs. We used the plsr.mcSimulation function to determine the VIP score and coefficients.

We also added a first attempt for the plot_pls() function (not final name obviously)


```{r}
# Apply plsr function to the mcSimulation object. Make sure the resultName has to be the name(character) of the column you want to perform the PLS

pls <- plsr.mcSimulation(object = test_mcSimulation_function,
                  resultName = names(test_mcSimulation_function$y)[3])


# Define the VIP function to avoid referring to chillR

VIP <- function(object) {
  
    if (object$method != "oscorespls")
      stop("Only implemented for orthogonal scores algorithm.  Refit with 'method = \"oscorespls\"'")
    if (nrow(object$Yloadings) > 1)
      stop("Only implemented for single-response models")
    
    SS <- c(object$Yloadings)^2 * colSums(object$scores^2)
    Wnorm2 <- colSums(object$loading.weights^2)
    SSW <- sweep(object$loading.weights^2, 2, SS / Wnorm2, "*")
    sqrt(nrow(SSW) * apply(SSW, 1, cumsum) / cumsum(SS))
  }

# Extract the VIP scores with VIP() from chillR. Not sure now what is "Comp 1" (component we guess)

vipResult <- VIP(pls)["Comp 1", ]

# The same for the pls coefficients. The use of ,, is new to me

coef <- pls$coefficients[,, 1]
      
# Create a df for further plotting

pls_outputs <- data.frame(Variable = names(vipResult),
                          VIP = vipResult,
                          Coef = coef)

# Remove the ugly rownames

rownames(pls_outputs) <- NULL


# Firt attempt of PLS plot

ggplot(filter(pls_outputs, VIP > 1),
       aes(VIP, reorder(Variable, VIP), fill = Coef > 0)) +
  geom_col() +
  geom_vline(aes(xintercept = 1))


```




## References
