---
title: "Applying the mcSimulation function in decisionSupport"
author: "Cory Whitney, Eduardo Fernandez, and Eike Luedeling <br /><small>INRES-Horticultural Sciences, University of Bonn, Auf dem Huegel 6, 53121 Bonn, Germany<small/>"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Applying the mcSimulation function in decisionSupport}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography:
 - packages.bib
 - vignettes.bib 
---


<style>
body {
text-align: justify}
</style>

```{r, include = F}
#set global options for knitr chunks 
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.width=5, 
  fig.height=3.5
)
```

```{r, warning = F, include = F}
#Automatically write R package citation entries to a .bib file
knitr::write_bib(c(.packages(), 
                   'chillR',
                   'dplyr',
                   'plyr',
                   'tidyverse',
                   'ggplot2', 
                   'decisionSupport'), 'packages.bib')

```

### Decision analysis with the decisionSupport package

#### The decisionSupport package

Many of the variables decision makers need to consider in development cannot be precisely quantified, at least not without unreasonable effort. The major objective of (prescriptive) decision analysis is to support the decision-making processes when faced with this problem [@Luedeling2016]. Decision analysis can make forecasts of decision outcomes without precise numbers. 

The decisionSupport package [@R-decisionSupport] implements this as a Monte Carlo simulation, which generates a large number of plausible system outcomes, based on random numbers for each input variable, drawn from user-specified probability distributions. 

1)	an R function that predicts decision outcomes based on the variables named in a separate data table. This R function is customized by the user to address a particular decision problem to provide the decision analysis model.  
2)	an input table (in .csv format) specifying the names and probability distributions for all variables used in the decision model. These distributions aim to represent the full range of possible values for each component of the model.  

These two inputs are provided as arguments to the `mcSimulation` function, which conducts a Monte Carlo analysis with repeated model runs based on probability distributions for all uncertain variables. The data table and model are customized to fit the particulars of a specific decision.

`decisionSupport` package can be installed from CRAN with the `install.packages()` function.

```{r, eval = FALSE}
install.packages("decisionSupport")
```

The package can be loaded with the `library()` function for further use in this example.

```{r}
library(decisionSupport)
```


#### Example – Reservoir Sedimentation in the Upper Volta River Basin of Burkina Faso

We showcase a partial example of a Monte-Carlo-based selection of a sedimentation management strategy for a reservoir in the Upper Volta River Basin of Burkina Faso [@Lanzanova2019]. These reservoirs have multiple benefits for rural communities and are important for food security and livelihoods. Sedimentation is a major impediment for the functioning of these reservoirs. The design of an efficient sedimentation management intervention requires assessment of multiple uncertain quantities and risks. Decision analysis is highly applicable in such data-scarce environments, where science has traditionally struggled to provide actionable information to policy-makers, development practitioners, NGOs and rural communities. 

Here we walk through a simple example of the use of various decisionSupport functions, which use the `tidyverse` libraries [@tidyverse2019] including `ggplot2` [@R-ggplot2], `plyr` [@R-plyr] and `dplyr` [@R-dplyr] among others in the [R programming language](https://www.r-project.org/) [@R-base]. 

Here we generate a model as a function using `decisionSupport` library. The example is a simple version of the model applied in Burkina Faso [@Lanzanova2019].  We use the `decisionSupport` functions `vv()` to produce time series with variation from a pre-defined mean and coefficient of variation, `chance_event()` to simulate whether events occur and `discount()` to discount values along a time series.

```{r model}
example_decision_function <- function(x, varnames){
  
  # calculate ex-ante risks: impact the implementation of interventions ####
  intervention_NonPopInvolvEvent <-
    chance_event(intervention_NonPopInvolv, 1, 0, n = 1)
  
  # pre-calculation of common random draws for all intervention model runs ####
  
  # profits from Tropical Livestock Units (TLU)
  TLU <- vv(TLU_no_intervention, var_CV, n_years)
  TLU_profit <- vv(profit_per_TLU, var_CV, n_years)
  
  # benefits of fruit
  precalc_intervention_fruit_benefits <-
    vv(intervention_fruit_area_ha, var_CV, n_years) *
    vv(intervention_fruit_yield_t_ha, var_CV, n_years) *
    vv(intervention_fruit_profit_USD_t, var_CV, n_years)
  
  # benefits of vegetables
  precalc_intervention_vegetable_benefits <-
    vv(intervention_vegetable_area_ha, var_CV, n_years) *
    vv(intervention_vegetable_yield_t_ha, var_CV, n_years) *
    vv(intervention_vegetable_profit_USD_t, var_CV, n_years)
  
  # benefits of rainfed crops
  precalc_intervention_rainfed_crop_benefits <-
    vv(intervention_rainfed_crop_area_ha, var_CV, n_years) *
    vv(intervention_rainfed_crop_yield_t_ha, var_CV, n_years) *
    vv(intervention_rainfed_crop_profit_USD_t, var_CV, n_years)
  
  #  Intervention ####
  
  for (decision_intervention_strips in c(FALSE,TRUE))
      {
      
  if (decision_intervention_strips)
  {
    intervention_strips <- TRUE
    intervention_strips_PlanningCost <- TRUE
    intervention_strips_cost <- TRUE
  } else
  {
    intervention_strips <- FALSE
    intervention_strips_PlanningCost <- FALSE
    intervention_strips_cost <- FALSE
  }
  
  if (intervention_NonPopInvolvEvent) {
    intervention_strips <- FALSE
    intervention_strips_cost <- FALSE
  }
  
  # Costs ####
  if (intervention_strips_cost) {
    cost_intervention_strips <-
      intervention_adaptation_cost + intervention_tech_devices_cost + intervention_nursery_cost +
      intervention_wells_cost +
      intervention_training_cost + intervention_mngmt_oprt_cost + intervention_mngmt_follow_cost +
      intervention_mngmt_audit_cost
  } else
    cost_intervention_strips <- 0
  
  if (intervention_strips_PlanningCost) {
    plan_cost_intervention_strips <-
      intervention_communication_cost + intervention_zoning_cost
  } else
    plan_cost_intervention_strips <- 0
  
  maintenance_cost <- rep(0, n_years)
  
  if (intervention_strips)
    maintenance_cost <-
    maintenance_cost + vv(maintenance_intervention_strips, var_CV, n_years)
  
  intervention_cost <- maintenance_cost
  intervention_cost[1] <-
    intervention_cost[1] + cost_intervention_strips + plan_cost_intervention_strips

  
  # Benefits from  cultivation in the intervention strips ####
  
  intervention_fruit_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_fruit_benefits
  intervention_vegetable_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_vegetable_benefits
  intervention_rainfed_crop_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_rainfed_crop_benefits
  
  # Total benefits from crop production (agricultural development and riparian zone) ####
  crop_production <-
    intervention_fruit_benefits +
    intervention_vegetable_benefits +
    intervention_rainfed_crop_benefits
  
  # Benefits from livestock ####
  # The following allows considering that intervention strips may
  # restrict access to the reservoir for livestock.
  
  if (intervention_strips)
    TLU_intervention <-
    TLU * (1 + change_TLU_intervention_perc / 100)
  else
    TLU_intervention <- TLU
  
  if (decision_intervention_strips){
    livestock_benefits <- TLU_intervention * TLU_profit
    total_benefits <- crop_production + livestock_benefits
    net_benefits <- total_benefits - intervention_cost
    result_interv <- net_benefits}
  
  
  if (!decision_intervention_strips){
    livestock_benefits <- TLU_no_intervention * TLU_profit
    total_benefits <- livestock_benefits
    net_benefits <- total_benefits - intervention_cost
    result_n_interv <- net_benefits}
  
    } #close intervention loop bracket

NPV_interv <-
  discount(result_interv, discount_rate, calculate_NPV = TRUE)

NPV_n_interv <-
  discount(result_n_interv, discount_rate, calculate_NPV = TRUE)

# Beware, if you do not name your outputs (left-hand side of the equal sign) in the return section, 
# the variables will be called output_1, _2, etc.

return(list(Interv_NPV = NPV_interv,
            NO_Interv_NPV = NPV_n_interv,
            NPV_decision_do = NPV_interv - NPV_n_interv,
            Cashflow_decision_do = result_interv - result_n_interv))
}

```

#### Perform a Monte Carlo simulation

Using the model function above, we can perform a Monte Carlo simulation with the `mcSimulation()` function from `decisionSupport`. This function generates distributions of all variables in the input table as well as the specified model outputs (see `return()` function above) by calculating random draws in our defined `example_decision_function()`. Make sure that all the variables in the input table are included in the model (erroneous variables listed there can cause issues with some of the post-hoc analyses). 

The `numberOfModelRuns` argument is an integer indicating the number of model runs for the Monte Carlo simulation. Unless the model function is very complex, 10,000 runs is a reasonable choice (for complex models, 10,000 model runs can take a while, so especially when the model is still under development, it often makes sense to use a lower number).

```{r mcSimulation}
test_mcSimulation_function <- decisionSupport::mcSimulation(
  estimate = decisionSupport::estimate_read_csv("example_input_table.csv"),
  model_function = example_decision_function,
  numberOfModelRuns = 1e3, #run 1,000 times
  functionSyntax = "plainNames"
)

```

#### Plot Net Present Value (NPV) distributions 

We can use the `plot_distributions()` function to produce one of the several plotting options for distribution outputs. This shows us an overlay of the full results of the Monte Carlo model of the decision options, i.e. the expected NPV if we choose to do the intervention `Interv_NPV` or not do the intervention `NO_Interv_NPV`.

```{r}
decisionSupport::plot_distributions(mcSimulation_object = test_mcSimulation_function, 
                                    vars = c("Interv_NPV", "NO_Interv_NPV"),
                                    method = 'smooth_simple_overlay')

```

We can use the same function for the value of the decision (difference in NPV between do and do not do). This is more helpful for us since it shows us the outcome distribution of the decision itself. 

```{r}
decisionSupport::plot_distributions(mcSimulation_object = test_mcSimulation_function, 
                                    vars = "NPV_decision_do",
                                    method = 'boxplot_density')

```

#### Cashflow analysis

Here we plot the distribution of annual cashflow over the entire simulated period for the intervention. For this we use the `plot_cashflow()` function which uses the specified cashflow outputs from the `mcSimulation()` function (in our case `Cashflow_decision_do`) to show cashflow over time.

```{r}
plot_cashflow(mcSimulation_object = test_mcSimulation_function, cashflow_var_name = "Cashflow_decision_do")

```

#### Projection to Latent Structures (PLS) analysis

We apply a post-hoc analysis to the `mcSimulation()` outputs with `plsr.mcSimulation()` to determine the Variable Importance in the Projection (VIP) score and coefficients of a Projection to Latent Structures (PLS) regression model. This function uses the outputs of the `mcSimulation()` selecting all the input variables from the decision analysis function in the parameter `object` and then runs a PLS regression with an outcome variable defined in the parameter `resultName`. We use the code `names(test_mcSimulation_function$y)[3]` to select the outcome variable `r names(test_mcSimulation_function$y)[3]`, which is the third element of the list `y` in our `test_mcSimulation_function` outputs (this must be a character element). 

```{r}
pls_result <- plsr.mcSimulation(object = test_mcSimulation_function,
                  resultName = names(test_mcSimulation_function$y)[3], ncomp = 1)

```

We run the `plot_pls()` on the results from `plsr.mcSimulation()` with a number of standard settings. The length of the bars is equal to VIP with a vertical line at '1' on the x-axis indicating a standard cut-off for VIP used for variable selection [@Whitney2017]. The overall plot only shows those variables with a VIP > 0.8, which is the common threshold for variable selection [@Lanzanova2019, @Luedeling2016]. The colors of the bars represent the positive or negative coefficient of the given input variable with the output variable.

Here we import the input table again to replace the labels for the variables on the y-axis. The input table can include a 'label' and 'variable' column. The standard labels (from the 'variable' column) are usually computer readable and not very nice for a plot. The `plot_pls()` function uses the text in the 'label' column as replacement for the default text in the 'variable' column.  

```{r}
input_table <- read.csv("example_input_table.csv")

plot_pls(pls_result, input_table = input_table)

```

#### Value of Information (VoI) analysis

We calculate Value of Information (VoI) analysis with the Expected Value of Perfect Information (EVPI). EVPI measures the expected opportunity loss that is incurred when the decision-maker does not have perfect information about a particular variable. EVPI is determined by examining the influence of that variable on the output value of a decision model. 

We use the function `data.frame()` to transform the x and y outputs of the `mcSimulation()` function for EVPI calculation. We use the ``multi_EVPI()` function to calculate the EVPI for multiple independent variables. For the first_out_var argument we choose 'intervention_mngmt_audit_cost' from the input table since this is the first variable after the NPV and cashflow model outputs, which we would like to exclude from the EVPI analysis.

```{r}
mcSimulation_table <- data.frame(test_mcSimulation_function$y, test_mcSimulation_function$x)

evpi <- multi_EVPI(mc = mcSimulation_table, first_out_var = "intervention_mngmt_audit_cost")

plot(x = evpi, out_var = "intervention_rainfed_crop_area_ha")
```


## References
